Asymmetric relationships and their effects on coexistence

### 2023-02-28 paper/Paper.Rnw
added a table for notation

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model is as follows.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table X.
%***DAN: Dan to make the table
For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

@@

%G4 ***DAN: Dan, this will have to be edited after Jasmin gets done editing the corresponding figures
Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition (Fig. \ref{fig:fig5}), and that 
ATA rescue occurs for many combinations of parameters for this system (Fig. \ref{fig:fig5},
but especially Fig. \ref{fig:fig6}).

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

@@

>>>>

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model is as follows.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

@@

%G4 ***DAN: Dan, this will have to be edited after Jasmin gets done editing the corresponding figures
Fulfilling goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition (Fig. \ref{fig:fig5}), and that 
ATA rescue occurs for many combinations of parameters for this system (Fig. \ref{fig:fig5},
but especially Fig. \ref{fig:fig6}).

@@

%***DAN: Jasmin, please add to the list anyone you talked to at UO about this, and anyone who read it and commented 
%for you. Be generous here - even short convos merit an acknowlegement.
The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Laurn Hallett, and 
Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

@@

\begin{table}[!h]
\caption{Summary of frequently used notation and abbreviations.}\label{tab:notation}
\begin{center}
\begin{tabular}{l|l}
Notation & Meaning \\
\hline 
\hline
\multicolumn{2}{c}{General} \\
\hline
ATA & Asymmetric tail associations \\
MCT & Modern coexistence theory \\
GWR & Growth rate when rare \\
E, C & Environment, competition \\
$\E$ & Expected value \\ 
ATA rescue & Coexistence occurs, but not after ATA effects are removed \\
ATA exclusion & Coexistence does not occur, but does after ATA effects are removed \\
 \hline
\multicolumn{2}{c}{Lottery model} \\
\hline
$N_i(t)$ & Population density of species $i$, time $t$ \\
$N$ & $N_1(t)+N_2(t)$ \\
$B_i(t)$ & Fecundity of species $i$, time $t$ \\
$\delta$ & Mortality rate \\
$\sigma$ & Standard deviation of log fecundities for the log-normal fecundities model \\
$\mu_i$ & Mean log fecundity, species $i$, log-normal fecundities model \\
$\eta_i$ & Upper bound fecundity, species $i$, log-normal fecundities model \\
\hline
\multicolumn{2}{c}{Theory} \\
\hline
$r_i(t)$ & Species $i$ growth rate, $\ln(N_i(t+1)/N_i(t))$ \\
$E_i(t)$ & Environment experienced by the $i^{\text{th}}$ species \\
$C_i(t)$ & Competition experienced by the $i^{\text{th}}$ species \\
$\olr_{i \bs i}$ & Mean GWR of species $i$ \\
$\olr_{j \bs i}$ & Mean growth of $j$ when $i$ is absent \\
$i \bs i$ & As subscript, refers to computing a quantity for $i$ when $i$ is rare \\
$j \bs i$ & As subscript, refers to computing a quantity for $j$ when $i$ is absent \\
$\shp$ & As superscript, refers to computing a quantity after rendering E and C independent \\
$\nat$ & As superscript, refers to computing a quantity after removing ATAs between E and C \\
$q_{ij}$ & Scaling factor \\
$\Delta_i^{(EC)}$ & Storage effects, species $i$ \\
$\Delta_i^{[EC]}$ & ATA effect to coexistence, species $i$ \\
$\Delta_i^{[E \nat C]}$ & Component of storage effects due to EC correlation \emph{per se}, species $i$ \\
\hline
\multicolumn{2}{c}{Diatom chemostat system} \\
\hline
$S$ & Extracellular silicate concentration \\
$x_1$ & Population density of \emph{F. crotonensis} \\
$x_2$ & Population density of \emph{C. pseudostelligera} \\
$\Theta$ & Temperature \\
$\Theta_0$ & Average temperature \\
$a$ & Amplitude of temperature fluctuations \\
$P$ & Period of temperature fluctuations \\
\hline
\end{tabular}
\end{center}
\end{table}
